---
pagetitle: "Code"
description-meta: "codechest — an R package of reusable functions for ANOVA workflows, multilevel models, correlations, factor analysis, reliability, and APA reporting, by Ryan T. Dobson."
page-layout: custom
toc: false
---

::: {.band .band-dark .band-tight}
::: {.wrap-narrow}
[Code]{.eyebrow}

## codechest {.section-title}

::: {.section-intro}
An R package of reusable functions for ANOVA workflows, multilevel models,
correlation tables, factor analysis, scale reliability, and APA-formatted
reporting.
:::

::: {.btn-row}
[View on GitHub](https://github.com/ryandobson/codechest){.btn-primary-c}
:::
:::
:::

::: {.band}
::: {.wrap-narrow}

::: {.prose}
**codechest** is an R package I began developing simply because I found myself repeatedly writing similar functions across different research projects. Over time, I started collecting and refining these functions so that I could move more efficiently between projects and maintain more consistent workflows.

The package is therefore best thought of as a **personal toolbox of reusable functions** for data cleaning, statistical modeling, and research workflows. I continue to expand and update the package as I encounter new problems that benefit from reusable solutions.

The code is still evolving and includes some quirks and idiosyncrasies that reflect the projects it grew out of. However, many of the functions may still be useful for others working on similar problems, and researchers are welcome to explore and adapt them.

Repository: [github.com/ryandobson/codechest](https://github.com/ryandobson/codechest)
:::

#### Install the package

You can install the development version of **codechest** from GitHub using:

```r
install.packages("remotes")
remotes::install_github("ryandobson/codechest")
```

```{r}
#| include: false

library(tidyverse)
library(codechest)

ll <- codechest::fivelovelanguages


```

:::
:::

::: {.band .band-cream}
::: {.wrap}

[What's inside]{.eyebrow}

## Function families {.section-title}

::: {.section-intro}
The package currently exports around ninety functions. They fall into a few
groups, most of which grew out of a specific analysis that I did not want to
write twice.
:::

::: {.card-grid}

::: {.c-card}
### ANOVA workflows

Assumption checks that interpret themselves, means tables, plots, a
multiverse of analysis specifications, and write-ups.

::: {.tags}
[anova_check()]{.tag}
[anova_workup()]{.tag}
[anova_multi()]{.tag}
:::
:::

::: {.c-card}
### Multilevel models

Within- and between-group decomposition, stepwise random-effects comparison,
formula construction, and fixed-effect trimming.

::: {.tags}
[mlm_groupmean()]{.tag}
[mlm_comparison()]{.tag}
[build_formulas()]{.tag}
:::
:::

::: {.c-card}
### Correlations

Publication-ready correlation matrices, including split tables that show two
groups above and below the diagonal.

::: {.tags}
[pub_cors_all()]{.tag}
[pub_cors_by()]{.tag}
[top_cor_diff()]{.tag}
:::
:::

::: {.c-card}
### Factor analysis and reliability

EFA summaries, factor loadings, scale reliability, and measurement invariance
testing across groups.

::: {.tags}
[efa_results()]{.tag}
[scale_reliabilities()]{.tag}
[test_measurement_invariance()]{.tag}
:::
:::

::: {.c-card}
### APA reporting

Model tables and reports formatted for APA style, with helpers to write them
straight out to Word.

::: {.tags}
[apa_lmer_model()]{.tag}
[apa_mlm_report()]{.tag}
[format_p()]{.tag}
:::
:::

::: {.c-card}
### Plotting and utilities

ggplot themes and palettes, quick distribution checks, and variable labels
that survive subsetting.

::: {.tags}
[dynamic_histogram()]{.tag}
[set_var_labels()]{.tag}
[gline_theme()]{.tag}
:::
:::

:::
:::
:::

::: {.band}
::: {.wrap}

[Examples]{.eyebrow}

## Example functions {.section-title}

::: {.section-intro}
Each example below runs against `fivelovelanguages`, a cleaned dataset from my
master's thesis that ships with the package.
:::

::: {.function-card}

### pub_cors_all()

`pub_cors_all()` quickly generates a publication-ready correlation table for a set of variables.


```{r}

#grab specific variable names as a vector
ll_aos <- names(ll[, grepl("rec_aos", colnames(ll))])

#create a vector to display names nicely (optional)
nice_names <- c(
  "oral sex"        = "rec_aos_pt_1",
  "pleasing sexual" = "rec_aos_pt_2"
)

pub_cors_all(data       = ll, #dataset
             cor_vars   = ll_aos,  #vector of variabl names from the dataset to include in analysis
             rename_vec = nice_names, #named vector to map new variable names
             cor_use    = "pairwise", #defaults to "pairwise" deletion of cases
             rnd        = 2 #default rounds to 2
             )

```


### pub_cors_by()


`pub_cors_by()` is an extension of pub_cors_all() that takes a dichotomous grouping variable and displays correlations for one group above the diagonal, and the other below.

I have updates to this function that I haven't incorporated into my package yet which includes:

- automatic calculation of the difference between grouping variable calculations (helps quickly identify interaction effects in correlations)

- easier argument specification so as automatically specify group levels based on the grouping variable provided



```{r}
pub_cors_by(data        = ll, #dataset
            cor_vars    = ll_aos,  #vector of variabl names from the dataset to include in analysis
            group_var   = "sex_bio",
            groups      = levels(ll$sex_bio),
            rename_vec  = nice_names, #named vector to map new variable names
            cor_use     = "pairwise", #defaults to "pairwise" deletion of cases
            rnd         = 2 #default rounds to 2
             )

```

:::


::: {.function-card}

### scale_reliabilities()

`scale_reliabilities()` computes reliability for several scales in one call, and `tidy_scale_reliabilities()` collapses the result into a single table you can drop straight into a manuscript.

Rather than just Cronbach's alpha, it also reports omega total and a unidimensionality index, which is useful when you want to know whether treating a set of items as one score is defensible in the first place.

```{r}
#| warning: false
#| message: false

# pull the item sets for four of the love languages (reception format)
grab <- function(pattern) grep(pattern, names(ll), value = TRUE)

love_language_scales <- list(
  "Words of affirmation" = grab("^rec_woa_[0-9]+$"),
  "Quality time"         = grab("^rec_qt_[0-9]+$"),
  "Physical touch"       = grab("^rec_pt_[0-9]+$"),
  "Gifts"                = grab("^rec_g_[0-9]+$")
)

rel <- scale_reliabilities(ll, scales = love_language_scales)

tidy_scale_reliabilities(rel, rnd = 2)
```

Passing a `group` argument returns the same table split by group, which makes it easy to check that a scale behaves similarly for everyone before you compare means across groups.

```{r}
#| warning: false
#| message: false

rel_by_sex <- scale_reliabilities(ll,
                                  scales = love_language_scales,
                                  group  = "sex_bio",
                                  groups = levels(ll$sex_bio))

tidy_scale_reliabilities(rel_by_sex, rnd = 2)
```

:::


::: {.function-card}

### mlm_groupmean()

`mlm_groupmean()` is a function that helps to easily decompose variables into within- and between-group variance. This is a really handy function if you do any multilevel model analyses.


```{r}
ll2 <- ll |> pivot_longer(cols      = c(rec_g_1, ex_g_1),
                              names_to  = "love_form",
                              values_to = "gift_1")
  mlm_groupmean(ll2,
              variables     = "gift_1",
              group         = "pid",
              within_affix  = "within_person_",
              between_affix = "between_person_",
              affix_type    = "prefix", #add a prefix. 'suffix' is the other option
              include_z     = TRUE #includes Z-scored versions of the variables
              ) |>
  select(matches("pid|gift_1|within_|between_")) |>
  arrange(pid) |>
  head()
```

:::

:::
:::
