---
title-block: false
toc: false
format: html
---

# Code

## codechest

**codechest** is an R package I began developing simply because I found myself repeatedly writing similar functions across different research projects. Over time, I started collecting and refining these functions so that I could move more efficiently between projects and maintain more consistent workflows.

The package is therefore best thought of as a **personal toolbox of reusable functions** for data cleaning, statistical modeling, and research workflows. I continue to expand and update the package as I encounter new problems that benefit from reusable solutions.

The code is still evolving and includes some quirks and idiosyncrasies that reflect the projects it grew out of. However, many of the functions may still be useful for others working on similar problems, and researchers are welcome to explore and adapt them.

Repository: [github.com/ryandobson/codechest](https://github.com/ryandobson/codechest)

#### Install the package

You can install the development version of **codechest** from GitHub using:

```r
install.packages("remotes")
remotes::install_github("ryandobson/codechest")
```


```{r}
#| label: setup
#| include: false

library(codechest)

ll <- codechest::fivelovelanguages


# nice_names <- c(
#   
# )



```


## Example functions

::: {.function-card}

### pub_cors_all()

`pub_cors_all()` quickly generates a publication-ready correlation table for a set of variables.


```{r}

#grab specific variable names as a vector
ll_aos <- names(ll[, grepl("rec_aos", colnames(ll))])

#create a vector to display names nicely (optional)
nice_names <- c(
  "oral sex"        = "rec_aos_pt_1",
  "pleasing sexual" = "rec_aos_pt_2"
)

pub_cors_all(data       = ll, #dataset 
             cor_vars   = ll_aos,  #vector of variabl names from the dataset to include in analysis
             rename_vec = nice_names, #named vector to map new variable names
             cor_use    = "pairwise", #defaults to "pairwise" deletion of cases
             rnd        = 2 #default rounds to 2
             )



```


`pub_cors_by()` is an extension of pub_cors_all() that takes a dichotomous grouping variable and displays correlations for one group above the diagonal, and the other below.

I have updates to this function that I haven't incorporated into my package yet which includes:
- automatic calculation of the difference between grouping variable calculations (helps quickly identify interaction effects in correlations)



```{r}

pub_cors_by(data        = ll, #dataset 
            cor_vars    = ll_aos,  #vector of variabl names from the dataset to include in analysis
            group_var   = "sex_bio",
            groups      = levels(ll$sex_bio),
            rename_vec  = nice_names, #named vector to map new variable names
            cor_use     = "pairwise", #defaults to "pairwise" deletion of cases
            rnd         = 2 #default rounds to 2
             )


```

